### 2015-03-14 manuscript/plos_template.tex
modified manuscript

Furthermore, both the signature 1 and 8 signatures shows strong abundance of T base at the $-2$ position,
and abundance of T bases at two bases 5' and 3' to the mutated site, respectively (Figures \ref{POLE1_sig_list} and \ref{POLE2_sig_list}).
In fact, nonsense mutations in TP53 at R231X, which are shown to be enriched in cancers having Pol $\epsilon$,
and abundance of T bases at two bases 5' and 3' to the mutated site, respectively (Figures \ref{POLE1_sig_list} and \ref{POLE2_sig_list}).
In the previous study of Pol $\epsilon$, a nonsense mutation in TP53 at R23X were shown to be enriched in cancers with Pol $\epsilon$ defects.
In fact, this mutation is C $>$ T at TpTpCpGpA, the most likely patten with the signature 8.
This implies that observing $\pm 2$ bases may be helpful for identifying the cause of mutations with higher resolution.

The signature 2 (C $>$ A at [CT]pCpT), observed in low grade gliomas, 
can be related to the one detected in the same cancer type in the previous study, but slightly different (see Supplementary Figure 7 (a)). 
This seems to be because the corresponding signature in the previous study shows very complex pattern (C $>$ A at NpCpT or C $>$ T at GpTpN).
Actually, the mutation related to this signature is mostly coming from one sample with a very high mutation rate (Supplementary Figure 7 (b,c)), 
and the signature 2 disappeared when we re-analyze the data removing this sample (Supplementary Figure 8).
We suspect that the complex low-grade-glioma specific signature detected in the previous study may be aggregation of multiple signatures,
but cannot be decomposed due to the lack of the number of samples having these signatures.
Therefore, we need to be more careful for the signature 2 and the corresponding signature in the previous study.

The signature 4 (C $>$ A at CpCpG) observed in kidney clear cell carcinomas, lung adenocarcinomas and melanomas 
shows remarkably high ratio of base C at the $-2$ position (see Supplementary Figure 9 (a,b,c)).
However, this signature is confide to only a few samples for every cancer type (see Supplementary Figure 9 (d, e, f)).
This seems to correspond to the ``signatures R2''  detected in kidney clear cell carcinomas, lung adenocarcinomas, lung squamous carcinomas and melanomas
(see supplementary figures of the previous study \cite{pmid23318258})
that could not be validated in the previous study,
meaning that significant ratios of somatic mutations corresponding to that signature was not estimated to be false positives by re-sequencing or visual inspection of BAM files using genomic viewers.
Actually, the pattern of this signature largely matches to that of the putative artifact by oxidation of DNA during acoustic shearing suggested in \cite{pmid23303777},
we guess that this signature is the same as what is observed in that study.
Although this signature may be from the artifacts, knowing higher resolution of the artifact signature is very helpful to accurately remove the false positive mutations.

The signature 13 (T $>$ [AGT] at TpCpN sites) was observed in 12 cancer types, and surely related to the activity of APOBEC family.
The composition of nucleotides in estimated signatures in 12 different cancer types are highly consistent (Figure \ref{APOBEC_sig_list}) except for the on in B-cell lymphoma,
and almost of of them showed slightly higher abundance of A and T bases and depletion of G base at the $-2$ position (Figure \ref{APOBEC_twofive}).
This fact backs up the results obtained from UCUT data in the previous subsection and the previous study \cite{pmid23318258},
and implies the robustness of the proposed approach.
The estimated transcribed strand-specificities varied across cancer types, implying that there is no significant strand-specificity in APOBEC signatures.
The Signature 15 and 16 may also be delated to the APOBEC signatures although the estimated forms are slightly different,
and are not merged to the cluster of signature 13 in the current criteria.

The signature 3 and 5 (C $>$ A at NpCpN) observed in head-and-neck cancers and three types of lung cancers are probably associated with tobacco smoking.
The estimated signature in each cancer type shows higher mutation prevalence on the template strand, which is consistent with the previous study \cite{pmid12379884, pmid23318258}.
The signature 6 (C $>$ A at NpCp[AT]) observed in neuroblastomas matches the pattern of that detected in the same cancer type in the previous study.
The signature 7 (C $>$ T at NpCpG sites) was observed in 25 out of 30 cancer type, and related to deamination of 5-methyl-cytosine.
The signature 9 (C $>$ T at NpCp[CT]) were observed in melanomas and glioblastomas, and are probably associated with a chemotherapy drug, temozolomide.
The signature 10 (C $>$T at [CT]pCpC) were observed in head and neck cancers and melanomas, and probably related to ultraviolet light.
The signature 11 (C $>$ T at GpCp[CG]) were observed in small-cell lung cancers and stomach cancers,
and seems to be the same with the ``signature 15'' in the previous study, whose function is still not clear,
and the depletion of C base at the $\pm 2$ position is consistently observed (Figure \ref{LMST_sig_list}).
the signature 12 (C $>$ T at [CG]pCp[CT]) observed in pilocytic astrocytomas,

The signature 18 (T $>$ C at ApTp[AG]) observed in liver cancer has been shown to be more common in Asian cases than in other ancestries \cite{pmid25362482}, 
though the source of this signature is still not clear.
The fact that very strong strains specificity is shown and 
the evidence of transcription-coupled nucleotide excision repair is implied is consistent to the previous studies \cite{pmid25362482, pmid23318258}.

In addition to the above signatures, a number of prominent mutation signatures observed in one cancer type 
had common characteristics with those observed in the previous study:
the signature 21 (T $>$ [CG] at CpTpT) observed in oesophagus cancers,

and the signature 19 (T $>$ C at GpTpN) in stomach cancers.

There results indicate that the proposed method can capture not only many of the prominent mutation signatures observed in the previous study,
but also novel characteristics for $\pm 2$ sites.

>>>>

Furthermore, both the signature 1 and 8 signatures shows strong abundance of T base at the $-2$ position,
and abundance of T bases at two bases 5' and 3' to the mutated site, respectively (Figures \ref{POLE1_sig_list} and \ref{POLE2_sig_list}, Supplementary Figure 7(b, c)).
In the previous study of Pol $\epsilon$, a nonsense mutation in TP53 at R23X were shown to be enriched in cancers with Pol $\epsilon$ defects.
In fact, this mutation is C $>$ T at TpTpCpGpA, the most likely patten with the signature 8.
This implies that observing $\pm 2$ bases may be helpful for identifying the cause of mutations with higher resolution.

The signature 2 (C $>$ A at [CT]pCpT), observed in low grade gliomas, 
can be related to the one detected in the same cancer type in the previous study, but slightly different (see Supplementary Figure 8 (a)). 
This seems to be because the corresponding signature in the previous study shows very complex pattern (C $>$ A at NpCpT or C $>$ T at GpTpN).
Actually, the mutation related to this signature is mostly coming from one sample with a very high mutation rate (Supplementary Figure 8 (b,c)), 
and the signature 2 disappeared when we re-analyze the data removing this sample (Supplementary Figure 9).
We suspect that the complex low-grade-glioma specific signature detected in the previous study may be aggregation of multiple signatures,
but cannot be decomposed due to the lack of the number of samples having these signatures.
Therefore, we need to be more careful for the signature 2 and the corresponding signature in the previous study.

The signature 4 (C $>$ A at CpCpG) observed in kidney clear cell carcinomas, lung adenocarcinomas and melanomas 
shows remarkably high ratio of base C at the $-2$ position (see Supplementary Figure 10 (a,b,c)).
However, this signature is confide to only a few samples for every cancer type (see Supplementary Figure 10 (d, e, f)).
This seems to correspond to the ``signatures R2''  detected in kidney clear cell carcinomas, lung adenocarcinomas, lung squamous carcinomas and melanomas
(see supplementary figures of the previous study \cite{pmid23318258})
that could not be validated in the previous study,
meaning that significant ratios of somatic mutations corresponding to that signature was not estimated to be false positives by re-sequencing or visual inspection of BAM files using genomic viewers.
Actually, the pattern of this signature largely matches to that of the putative artifact by oxidation of DNA during acoustic shearing suggested in \cite{pmid23303777},
we guess that this signature is the same as what is observed in that study.
Although this signature may be from the artifacts, knowing higher resolution of the artifact signature is very helpful to accurately remove the false positive mutations.

The signature 13 (T $>$ [AGT] at TpCpN sites) was observed in 12 cancer types, and surely related to the activity of APOBEC family.
The composition of nucleotides in estimated signatures in 12 different cancer types are highly consistent (Figure \ref{APOBEC_sig_list}) except for the on in B-cell lymphoma,
and almost of of them showed slightly higher abundance of A and T bases and depletion of G base at the $-2$ position (Figure \ref{APOBEC_twofive}, Supplementary Figure 7(a)).
This fact backs up the results obtained from UCUT data in the previous subsection and the previous study \cite{pmid23318258},
and implies the robustness of the proposed approach.
The estimated transcribed strand-specificities varied across cancer types, implying that there is no significant strand-specificity in APOBEC signatures.
The Signature 15 and 16 may also be delated to the APOBEC signatures although the estimated forms are slightly different,
and are not merged to the cluster of signature 13 in the current criteria.

The signature 3 and 5 (C $>$ A at NpCpN) observed in head-and-neck cancers and three types of lung cancers are probably associated with tobacco smoking.
The estimated signature in each cancer type shows higher mutation prevalence on the template strand, which is consistent with the previous study \cite{pmid12379884, pmid23318258}.
The signature 6 (C $>$ A at NpCp[AT]) observed in neuroblastomas matches the pattern of that detected in the same cancer type in the previous study.
The signature 7 (C $>$ T at NpCpG sites) was observed in 25 out of 30 cancer type, and related to deamination of 5-methyl-cytosine.
The signature 9 (C $>$ T at NpCp[CT]) were observed in melanomas and glioblastomas, and are probably associated with a chemotherapy drug, temozolomide.
The signature 10 (C $>$T at [CT]pCpC) were observed in head and neck cancers and melanomas, and probably related to ultraviolet light.
Observed consistent strand specificities (Figure \ref{UV_sig_list}) matches to the previous results \cite{pmid23318258}.   
Also, higher frequencies of T base at the $\pm 2$ position are consistently observed (Figure \ref{UV_sig_list}, Supplementary Figure 7(d)).
The signature 11 (C $>$ T at GpCp[CG]) were observed in small-cell lung cancers and stomach cancers,
and seems to be the same with the ``signature 15'' in the previous study, whose function is still not clear,
and the abundance of G base at the $\pm 2$ position is consistently observed (Figure \ref{LMST_sig_list}, Supplementary Figure 7(e)).

The signature 18 (T $>$ C at ApTp[AG]) observed in liver cancer has been shown to be more common in Asian cases than in other ancestries \cite{pmid25362482}, 
though the source of this signature is still not clear. 
In this signature, we can observe very strong strains specificity implying the evidence of transcription-coupled nucleotide excision repair as shown in \cite{pmid25362482, pmid23318258}.
The unknown signatures 12 (C $>$ T at [CG]pCp[CT]), 19 (T $>$ C at GpTpN) and 21 (T $>$ [CG] at CpTpT) observed in pilocytic astrocytomas, stomach cancers and oesophagus cancers, respectively,
coincides well with the those detected in the same cancer types in the previous study \cite{pmid23318258}.
Although several interesting characteristics were observed at the $\pm$ positions for these signatures, 
their validity are difficult to confirm at this point since these are not reproduced in multiple cancer types.

In summary, the propose method can capture many of the prominent mutation signatures observed in the previous study.
Also, the propose method detected several novel characteristics at the $\pm 2$ position that is consistent across multiple cancer types,
suggesting the existence of more hidden characteristics at the distant position from the mutated site.

### 2015-03-14 supp/supp.tex
modified manuscript

% \subfigure[POLE1 signatures in each cancer type]{%
%  \includegraphics[width=7.5cm,height=2.5cm]{signatureList_POLE1.eps}
%   \label{POLE1_sig_list}}

% \subfigure[POLE2 signature in each cancer type]{%
%   \includegraphics[width=7.5cm,height=2.5cm]{signatureList_POLE2.eps}
%   \label{POLE2_sig_list}}

% \subfigure[UV signature in each cancer type]{%
%   \includegraphics[width=7.5cm,height=2.5cm]{signatureList_UV.eps}
%   \label{UV_sig_list}}

% \subfigure[Unknown signature obtained in lung small cell carcinomas and stomach cancers]{%
%   \includegraphics[width=7.5cm,height=2.5cm]{signatureList_LMST.eps}
%   \label{LMST_sig_list}}

% \caption{Several signatures having prominent characteristics at 5'  or 3' to the mutated sites.}
% \label{nature2013_sig_others}
% \end{figure*}

>>>>

\subfigure[APOBEC signature intensities at two 5' to the mutated site]{%
  \includegraphics[width=14.5cm,height=6cm]{APOBEC_two5prime.eps}
   \label{APOBEC_two5prime}}

\subfigure[POLE1 signature intensities at two 5' to the mutated site]{%
  \includegraphics[width=4.5cm,height=6cm]{POLE1_two5prime.eps}
  \label{POLE1_two5prime}}
\quad
\subfigure[POLE2 signature intensities at two 5' to the mutated site]{%
  \includegraphics[width=4.5cm,height=6cm]{POLE2_two5prime.eps}
  \label{POLE2_two5prime}}

\subfigure[UV signature intensities at two 5' to the mutated site]{%
  \includegraphics[width=4.5cm,height=6cm]{UV_two5prime.eps}
  \label{UV_two5prime}}
\quad
\subfigure[The signature 11 intensities at two 5' to the mutated site]{%
  \includegraphics[width=4.5cm,height=6cm]{LMST_two5prime.eps}
  \label{LMST_two5prime}}

\caption{The estimated frequencies of bases at two 5' to the mutated site for each cancer type.
The heights of bar show the estimated frequency for bases A, C, G and T at two 5' to the mutated site.
The error bars show standard errors estimated by bootstrap.}
\label{two5prime}

\end{figure*}
